---
title: "Project 1"
author: "Tatev Gomtsyan"
date: "2023-06-26"
output: html_document
---

```{r}
diamonds <- read.csv("C:/Users/Student/Desktop/MSDS/Summer '23/STAT 6021/Project 1/diamonds4.csv")
head(diamonds)
```
```{r}
# cut vs price association 

#reorder bars to make more sense 
diamonds$cut = factor(diamonds$cut,levels = c("Good","Very Good", "Ideal", "Astor Ideal"))

new<- diamonds %>% 
  group_by(cut) %>% 
  summarize(count=n()) %>% 
  mutate(percentage = round(count / sum(count) * 100, 2))
new

#bar graph
cut <- ggplot(new, aes(x=cut, y=percentage))+
  geom_bar(stat="identity", fill="lightblue")+
  labs(x="Cut Type", y="Proportion of Purchases", title="Proportion of Purchases by Cut")+
 # scale_fill_manual(values = c("red", "blue", "orange", "green"))+
  theme_minimal()
cut
#scale price

```
The most popular cut type among buyers is "Ideal" as it has the highest proportion. Astor Ideal has the lowest number of purchases from Blue Nile, most likely due to its rarity. We will see later that price does not influence the number of purchases of the Astor Ideal, as it is has the lowest price range and smallest carat number out of all the cuts. 

```{r}
ggplot(diamonds, aes(x=carat, y=price, color=cut)) +
    geom_point() +
    stat_smooth(method='lm', size=0.5)+ #adding regression lines
    labs(x="Carat", y = "Price", title="How Price Varies by Cut Type and Carat")+
    theme(plot.title = element_text(hjust = 0.5))
```
```{r}
options(scipen = 999)#gets rid of scientific notation
library(scales)
ggplot(diamonds, aes(x=carat, y=price, color=cut)) +
    geom_point() +
    stat_smooth(method='lm', size=0.5)+ #adding regression lines
    labs(x="Carat", y = "Price", title="How Price Varies by Cut Type and Carat")+
    facet_wrap(~cut)+
    theme(plot.title = element_text(hjust = 0.5))
```

As carat increases, price also increases for all diamond cut types. By observation, Good cut has the steepest slope. The highest prices belong to Very Good and Ideal cuts. Astor Ideal cuts usually have smaller carats and lower price. This suggests that carat size has a large influence on the price point, more so than the cut itself. 

```{r}
med_price<- diamonds %>% 
  group_by(cut) %>% 
  summarise(medPrice=median(price))
med_price
```

```{r}
avg_price<- diamonds %>% 
  group_by(cut) %>% 
  summarise(avgPrice=mean(price))
avg_price
```

```{r}
log_price<- log(diamonds$price) #restores symmetry to data and reduces skewness
ggplot(diamonds, aes(x=cut, y=log_price)) +
  geom_boxplot(fill="pink")+
  labs(x="Cut Type", y="Price", title="Boxplot Representing Price vs Cut of Blue Nile Diamonds")+
  theme(plot.title = element_text(hjust = 0.5)) #centering title
```
We log transformed the price to create a more visually effective boxplot. Judging by the figure, we can see that the IQR of prices of all cut types is about 7-9. Astor Ideal has the highest median out of the cut types. Good and Very Good cuts have almost identical medians, but the Very Good cut has a larger range of prices and a much higher maximum price. Surprisingly, Ideal cuts have the biggest IQR and the first quartile of this box has the lowest starting point. 

```{r}
cut_color <- diamonds %>%
  group_by(cut, color) %>%
  summarize(meanPrice=mean(price))
cut_color
```

```{r}
ggplot(diamonds, aes(x=cut, fill=color))+
  geom_bar(position="fill")+
  labs(y="proportion", title="Cut by Color")
```
The highest quality color is considered 'D', and the lowest, 'J'. Overall, there doesn't seem to be too much difference in the colors of each cut. However, we can make the observation that Good cut has the largest proportion of D color, while Astor Ideal has the largest prop of F color. When thinking about the price points of these cuts and the quality of colors, we know that Astor Ideal doesn't have a high price point, which is in line with the color quality. On the other hand, we know that Very Good and Ideal cuts have the highest prices, but this is not reflected in the colors that are used for these cuts. All colors seem to be used pretty eveny for both cuts, with no overwhelming use of any color to create an association with price. 

